---
title: "ForestFirePredictionProject"
author: "Ziming Yang"
date: "12/9/2019"
output:
  word_document: default
  pdf_document: default
---

```{r cars}
library(corrplot)
library(car)
library(perturb)
library(ggplot2)
library(stargazer)
```

## Factor
```{r pressure, echo=FALSE}
forest <- read.csv("forestfires.csv")
mat1 <- model.matrix(~factor(month),data=forest)

# data.frame(forest$month, mat1)
```

# create a binary column for area, area with nonzero value is set to be 1, otherwise is 0.
```{r}
forest$area_binary <- ifelse(forest$area==0, 0, 1)
forest$area_binary
forest$area
forest <- data.frame(forest)
```

# carete a new dataset with only observations of burned area greater than 0. Name this new dataset as forest_2
```{r}
forest_2 <- forest[forest$area>0,]
forest_2$area
nrow(forest_2)
nrow(forest)
nrow(forest[forest$area==0,])
```

# log transformation on some variables
```{r}
forest$log_rain <- log(forest$rain)
forest$log_FFMC <- log(forest$FFMC)
forest$log_ISI <- log(forest$ISI)
forest_2$log_area <- log(forest_2$area)
hist(forest_2$log_area)
```

## logistic regression
```{r}

```



# plot histograms for variables in dataset forest
```{r}



hist(forest$DC)
png(filename = "DC_hist.png", width = 3000, height = 3000, res = 300)
par(cex = 1.3, mar = c(5, 4, 2, 1))
hist(forest$DC,
breaks = "FD",
xlab = "DC",
main = "")
dev.off()


hist(forest$log_ISI,col="gray")
png(filename = "log_ISI_hist.png", width = 3000, height = 3000, res = 300)
par(cex = 1.3, mar = c(5, 4, 2, 1))
hist(forest$log_ISI,col="green",
breaks = "FD",
xlab = "log_ISI",
main = "")
dev.off()

hist(forest_2$log_area,col="gray")
png(filename = "log_area_hist.png", width = 3000, height = 3000, res = 300)
par(cex = 1.3, mar = c(5, 4, 2, 1))
hist(forest_2$log_area,col="gray",
breaks = "FD",
xlab = "log_area",
main = "log(area)")
dev.off()
```



Bar plot
```{r}
# binary
png(filename = "binary_area.png", width = 3000, height = 3000, res = 300)
#par(cex = 1.3, mar = c(5, 4, 2, 1))
barplot(prop.table(table(forest$area_binary)), beside=TRUE, space=c(2, 3))
dev.off()
barplot(prop.table(table(forest$area_binary)))
# month
png(filename = "month_freq.png", width = 3000, height = 3000, res = 300)
par(cex = 1.3, mar = c(5, 4, 2, 1))
barplot(prop.table(table(forest$month)))
dev.off()
barplot(prop.table(table(forest$month)))
# day
png(filename = "day_freq.png", width = 3000, height = 3000, res = 300)
par(cex = 1.3, mar = c(5, 4, 2, 1))
barplot(prop.table(table(forest$day)))
dev.off()
barplot(prop.table(table(forest$day)))
```

# build model
```{r}
# model 1 linear regression with burned area greater than zero
model_1 <- lm(log(area)~X+Y+factor(day)+FFMC+DMC+DC+ISI+temp+RH+wind+rain,data=forest_2)

# stargazer(model_1, model_2,model_3, header=FALSE,type = "text")
# plot fitted value
## model 1
plot(fitted(model_1))
lines(lowess(fitted(model_1),rstudent(model_1)), col = "green", lwd = 4)
png(filename = "model1_fitted.png", width = 3000, height = 3000, res = 300)
par(cex = 1.5, mar = c(5, 4, 2, 1))
plot(fitted(model_1), rstudent(model_1),
     xlab = "Fitted Values",
     ylab = "Studentized Residuals",
     pch  = 19,
     col  = rgb(0, 0, 0, .4))
lines(lowess(fitted(model_1),rstudent(model_1)), col = "green",lwd = 4)
dev.off()
## mdoel 2
model_2 <- lm(log(area)~X+Y+factor(day)+log(FFMC)+DMC+DC+ISI+temp+RH+wind+rain,data=forest_2)
png(filename = "model2_fitted.png", width = 3000, height = 3000, res = 300)
par(cex = 1.5, mar = c(5, 4, 2, 1))
plot(fitted(model_2), rstudent(model_2),
     xlab = "Fitted Values",
     ylab = "Studentized Residuals",
     pch  = 19,
     col  = rgb(0, 0, 0, .4))
lines(lowess(fitted(model_2),rstudent(model_2)), col = "green",lwd = 4)
dev.off()
## model 3

model_3 <- lm(log(area)~X+Y+factor(day)+log(FFMC)+DMC+DC+log(ISI)+temp+RH+wind+rain,data=forest_2)
plot(fitted(model_3))
png(filename = "model3_fitted.png", width = 3000, height = 3000, res = 300)
par(cex = 1.5, mar = c(5, 4, 2, 1))
plot(fitted(model_3), rstudent(model_3),
     xlab = "Fitted Values",
     ylab = "Studentized Residuals",
     pch  = 19,
     col  = rgb(0, 0, 0, .4))
lines(lowess(fitted(model_3),rstudent(model_3)), col = "green",lwd = 4)
dev.off()
```
```{r}
round(summary(model_3)$coefficients,3)
plot_coeffs <- function(mlr_model) {
  coeffs <- coefficients(mlr_model)
  mp <- barplot(coeffs, col="grey", xaxt='n', main="Regression Coefficients")##3F97D0
  lablist <- names(coeffs)
  text(mp, par("usr")[3], labels = lablist, srt = 45, adj = c(1.1,1.1), xpd = TRUE, cex=0.6)
}
plot_coeffs(model_3)
```


## fitted value
```{r}
plot(fitted(model_1), rstudent(model_1))
png(filename = "fitted.png", width = 3000, height = 3000, res = 300)
par(cex = 1.5, mar = c(5, 4, 2, 1))
plot(fitted(model_1), rstudent(model_1),
xlab = "Fitted Values",
ylab = "Studentized Residuals",
ylim = c(-4, 4),
xlim = c(20000, 160000),
pch = 19,
col = rgb(0, 0, 0, .4))
lines(lowess(fitted(model_1),rstudent(model_1)), col = "green", lwd = 4)
```

# GAM
```{r}
library(mgcv)
# model 4
model_4 <- gam(log(area)~s(X,k=6)+Y+factor(month)+factor(day)+log(FFMC)+DMC+DC+ISI+temp+RH+wind+rain,data=forest_2)
png(filename = "model4_gam.png", width = 3000, height = 3000, res = 300)
par(cex = 1.5, mar = c(5, 4, 2, 1))
plot(model_4, 
     residuals = TRUE, 
     shade    = TRUE,
     pch = 19,
     col = rgb(0, 0, 0, .4),
     cex = .5)
dev.off()

forest_2$X_plus_Y <- forest_2$X + forest_2$Y
# model 5
model_5 <- gam(log(area)~s(X_plus_Y,k=6)+factor(month)+factor(day)+log(FFMC)+DMC+DC+ISI+temp+RH+wind+rain,data=forest_2)
png(filename = "model5_gam.png", width = 3000, height = 3000, res = 300)
par(cex = 1.5, mar = c(5, 4, 2, 1))
plot(model_5, 
     residuals = TRUE, 
     shade    = TRUE,
     pch = 19,
     col = rgb(0, 0, 0, .4),
     cex = .5)
dev.off()

```
# cross validation
```{r}
# generate empty prediction and outcome vectors
predictions1 <- c()
predictions2 <- c()
predictions3 <- c()
predictions4 <- c()
predictions5 <- c()


# logistic regression CV function
CV <- function(data1 = forest_2, k = 5){
  # select number of folds
  folds <- k
  # generate fold sequence
  fold.num <- rep(1:folds, length.out = nrow(data1))
  # randomize fold sequence
  fold.samp <- sample(fold.num)
  for(i in 1:k){
    
    # Training data 
    train <- data1[fold.samp != i, ]
    # Test data takes the remaining rows
    test <- data1[fold.samp == i, ]
    
    # Run lm on training data
    model_1 <- lm(log(area)~X+Y+day+FFMC+DMC+DC+ISI+temp+RH+wind+rain,data=train)
    model_2 <- lm(log(area)~X+Y+day+log(FFMC)+DMC+DC+ISI+temp+RH+wind+rain,data=train)
    model_3 <- lm(log(area)~X+Y+day+log(FFMC)+DMC+DC+log(ISI)+temp+RH+wind+rain,data=train)
    model_4 <- gam(log(area)~s(X,k=5)+s(Y,k=5)+day+log(FFMC)+DMC+DC+ISI+temp+RH+wind+rain,data=train)
    model_5 <- gam(log(area)~s(X_plus_Y,k=5)+day+log(FFMC)+DMC+DC+ISI+temp+RH+wind+rain,data=train)
    
    # Test error
    model_1.test <- (log(test$area) - predict(model_1,  test, type="response"))^2
    model_2.test <- (log(test$area) - predict(model_2,  test, type="response"))^2
    model_3.test <- (log(test$area) - predict(model_3,  test, type="response"))^2
    model_4.test <- (log(test$area) - predict(model_4,  test, type="response"))^2
    model_5.test <- (log(test$area) - predict(model_5, test, type="response"))^2

    # Store results
    options(warn=-1)
    predictions1 <- append(predictions1, model_1.test)
    predictions2 <- append(predictions2, model_2.test)
    predictions3 <- append(predictions3, model_3.test)
    predictions4 <- append(predictions4, model_4.test)
    predictions5 <- append(predictions5, model_5.test)


    # Test rmse
    model_1.rMSE <- sqrt(sum(predictions1)/(length(predictions1)))
    model_2.rMSE <- sqrt(sum(predictions2)/(length(predictions2)))
    model_3.rMSE <- sqrt(sum(predictions3)/(length(predictions3)))
    model_4.rMSE <- sqrt(sum(predictions4)/(length(predictions4)))
    model_5.rMSE <- sqrt(sum(predictions5)/(length(predictions5)))
    }
    return(list(model_1.rMSE, model_2.rMSE, model_3.rMSE,model_4.rMSE,model_5.rMSE))
}

# Replicate the 5-fold cross-validation 100 times
cvs <- replicate(100, CV())

# Pull out the average rMSE for each model
model_1.cv <- mean(sapply(cvs[1, ], mean))
model_2.cv <- mean(sapply(cvs[2, ], mean))
model_3.cv<- mean(sapply(cvs[3, ], mean))
model_4.cv<- mean(sapply(cvs[4, ], mean))
model_5.cv<- mean(sapply(cvs[5, ], mean))
res <- c(model_1.cv,model_2.cv,model_3.cv,model_4.cv,model_5.cv)
res
```

# logistic regression
```{r}
model_6 <- glm(area_binary ~ X+Y+day+log(FFMC)+DMC+DC+log(ISI+1)++temp+RH+wind+rain,
data = forest, family = binomial(link = "logit"))
```

```{r}
round(summary(model_6)$coefficients,3)
plot_coeffs <- function(mlr_model) {
  coeffs <- coefficients(mlr_model)
  mp <- barplot(coeffs, col="grey", xaxt='n', main="Regression Coefficients")##3F97D0
  lablist <- names(coeffs)
  text(mp, par("usr")[3], labels = lablist, srt = 45, adj = c(1.1,1.1), xpd = TRUE, cex=0.6)
}
plot_coeffs(model_6)
```

```{r}
newdata
```

```{r}
Wed <- subset(forest, day =="wed")
Tue <- subset(forest, day =="tue")
Sat <- subset(forest, day =="sat")
Sun <- subset(forest, day =="sun")
Fri <- subset(forest, day =="fri")
png(filename = "wed.png", width = 3000, height = 3000, res = 300)
barplot(prop.table(table(Wed$area_binary)), beside=TRUE, space=c(2, 3),main="Wednesday")
dev.off()
png(filename = "tue.png", width = 3000, height = 3000, res = 300)
barplot(prop.table(table(Tue$area_binary)), beside=TRUE, space=c(2, 3),main="Tuesday")
dev.off()
png(filename = "sat.png", width = 3000, height = 3000, res = 300)
barplot(prop.table(table(Sat$area_binary)), beside=TRUE, space=c(2, 3),main="Saturday")
dev.off()
png(filename = "sun.png", width = 3000, height = 3000, res = 300)
barplot(prop.table(table(Sun$area_binary)), beside=TRUE, space=c(2, 3),main="Sunday")
dev.off()
barplot(prop.table(table(Fri$area_binary)), beside=TRUE, space=c(2, 3),main="Sunday")

```

```{r}
barplot(prop.table(table(Fri$area_binary)), beside=TRUE, space=c(2, 3),main="Sunday")
```

